---
title: "Run initial two-way ANOVAS"
subtitle: "Quasi Dark-Adapted Rapid Light Curve (RLC) Analysis"
author: "Sarah Tanja"
date: "`r format(Sys.time(), '%d %B, %Y')`"  
format:
  html:
    df-print: paged
    toc: true
    toc-location: right
    smooth-scroll: true
    link-external-icon: true
    link-external-newwindow: true
    code-fold: false
    code-tools: true
    code-copy: true
    highlight-style: breeze
    code-overflow: wrap
    theme: minty
editor: 
  markdown: 
    wrap: 72
---

# Overview

Quasi dark-adapted rapid light curve analysis from measurements taken by
a Walz Diving PAMII...

# Install and load packages

```{r}
# Install packages
if ("tidyverse" %in% rownames(installed.packages()) == 'FALSE') install.packages('tidyverse')
if ("dplyr" %in% rownames(installed.packages()) == 'FALSE') install.packages('dplyr')
if ("lubridate" %in% rownames(installed.packages()) == 'FALSE') install.packages('lubridate')
if ("stringr" %in% rownames(installed.packages()) == 'FALSE') install.packages('stringr')

# Load packages
library(dplyr)
library(tidyverse)
library(lubridate)
library(stringr)
```

# Pull in data

This data is generated by the report tab in the Win-3 Control Software
of the Walz Diving PAM II. Of importance to note is that regression 1
(REG1) and regression 2 (REG2) for each light curve are listed above the
tabular light curve data.

```{r}
rlc_0619 <- read_csv("../pam-data/rlc_06192024.csv")
rlc_0620 <- read_csv("../pam-data/rlc_06202024.csv")
rlc_0621 <- read_csv("../pam-data/rlc_06212024.csv")
rlc_0622 <- read_csv("../pam-data/rlc_06222024.csv")
```


# FV/Fm

```{r}
fvfm_aov <- aov(FvFm ~ treatment,
  data = photobot_results
)

summary(fvfm_aov)
```

```{r}

FvFm <- ggplot(photobot_results) +
  aes(x = treatment, y = FvFm, fill= treatment) +
  geom_boxplot() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
   labs(
    title = "Concentration vs. Photosynthetic efficiency (Fv/Fm)",
    x = "Leachate concentration (mg/L)",
    y = "Photosynthetic efficiency (Fv/Fm)"
  ) +
  scale_x_discrete(labels = c("Control ", "PPa 0.01", "PPa 0.1","PPa 1", "PPa 10", "PPa 100","PVC 0.01", "PVC 0.1", "PVC 1", "PVC 10", "PVC 100")) +  
  scale_fill_manual(values = c("seashell", "coral", "coral1","coral2","coral3", "coral4","dodgerblue","dodgerblue1","dodgerblue2", "dodgerblue3","dodgerblue4")) 

FvFm

# Save the ggplot object to a PNG file
ggsave(filename = "FvFm.png", plot = FvFm, width = 11, height = 6, dpi = 600)

```

# Alpha

## Day 1

```{r}
Alpha_1 <- ggplot(regression_0619) +
  aes(x = treatment, y = alpha, fill= leachate) +
  geom_boxplot() +
  scale_x_discrete(limits = c("ambient_control", "hot_control", 
                              "ambient_low", "hot_low", 
                              "ambient_high", "hot_high")) +
  scale_y_continuous(breaks = pretty(regression_0619$alpha, n = 5)) + # Limit number of ticks
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
   labs(
    title = "Alpha values on the 1st day of exposure",
    x = "",
    y = "Alpha"
  ) 

Alpha_1

# Save the ggplot object to a PNG file
#ggsave(filename = "Alpha.png", plot = Alpha, width = 11, height = 6, dpi = 300)
```

## anova

temp

```{r}
alpha_aov <- aov(alpha ~ temp,
  data = regression_0619
)

summary(alpha_aov)
```

leachate

```{r}
alpha_aov <- aov(alpha ~ leachate,
  data = regression_0619
)

summary(alpha_aov)
```

temp\*leachate

```{r}
alpha_aov <- aov(alpha ~ temp*leachate,
  data = regression_0619
)

summary(alpha_aov)
```

### plot

```{r}
PVCAlpha <- ggplot(photobot_pvc) +
  aes(x = treatment, y = alpha, fill= treatment) +
  geom_boxplot() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
   labs(
    title = "Concentration vs. Alpha",
    x = "Leachate concentration (mg/L)",
    y = "Alpha"
  ) +
  scale_x_discrete(labels = c("Control ", "PVC 0.01", "PVC 0.1", "PVC 1", "PVC 10", "PVC 100")) +  
  scale_fill_manual(values = c("seashell", "dodgerblue","dodgerblue1","dodgerblue2", "dodgerblue3","dodgerblue4")) 

PVCAlpha

# Save the ggplot object to a PNG file
ggsave(filename = "PVCAlpha.png", plot = PVCAlpha, width = 11, height = 6, dpi = 300)

```

Alpha (α) Definition: Alpha (α) is defined as the initial slope of the
Rapid Light Curve at low light intensities. It represents the
photosynthetic efficiency, indicating how effectively a photosynthetic
organism can use available light for photosynthesis under low light
conditions. A higher alpha value indicates that the photosynthetic
organism is highly efficient at utilizing low light levels, which can be
crucial for survival in shaded or low-light environments. Conversely, a
lower alpha suggests less efficiency in low light conditions. The alpha
value is calculated from the initial linear portion of the Rapid Light
Curve. It is determined by fitting a straight line to the data points at
the lowest light intensities and calculating the slope of this line. The
steeper the slope, the higher the photosynthetic efficiency at low
light.

### anova

```{r}
alpha_pvc_aov <- aov(alpha ~ treatment,
  data = photobot_pvc
)

summary(alpha_pvc_aov)
```

### Dunnett's test

Is the lowest PVC leachate concentration actually larger than the
control? Or is this a `Type II error`, or error of multiple tests? I
followed the [stats and r blog to conduct a post-hoc Dunnett's
test](https://statsandr.com/blog/anova-in-r/#post-hoc-test:~:text=plot(TukeyHSD(res_aov))-,Dunnett%E2%80%99s%20test,-We%20have%20seen)
on the above ANOVA result.

```{r}
if ("multcomp" %in% rownames(installed.packages()) == 'FALSE') install.packages('multcomp')
library(multcomp)

# Make sure `treatment` is a factor


# Dunnett's test:
post_test <- glht(alpha_pvc_aov,
  linfct = mcp(treatment = "Dunnett")
)

summary(post_test)
```

::: callout-warning
The Dunnett's test adjusted *p* value comparing the lowest PVC
concentration to the control (*p* = .0671) indicated that the treatment
is not significantly different from the control.
:::

::: callout-note
Note that in R, by default, the reference category for a [factor
variable](https://statsandr.com/blog/data-types-in-r/#factor) is the
first category in alphabetical order. This is the reason that, by
default, the reference species is `Control`.

The reference category can be changed with the `relevel()` function (or
with the [`{questionr}`
addin](https://statsandr.com/blog/rstudio-addins-or-how-to-make-your-coding-life-easier/#reordering-factors)).
Look
[here](https://statsandr.com/blog/anova-in-r/#post-hoc-test:~:text=%23%20Change%20reference%20category%3A)
for an example.
:::

Examining the 95% family-wise confidence level comparing the treatments
to the controls:

```{r}

# Open the PNG device
png(filename = "Dunnett.png", width = 800, height = 600, res = 72)

### use sufficiently large upper margin
old.par <- par(mai=c(1,2.25,1.25,0.25), no.readonly=TRUE)
### plot
plot(post_test)
par(old.par)

# Close the device
dev.off()

```

```{r}
### use sufficiently large upper margin
old.par <- par(mai=c(1,2.25,1.25,0.25))
### plot
plot(post_test)
par(old.par)
```

## PP only?

```{r}
photobot_pp <- photobot_results %>% 
  filter(treatment %in% c('Control', 'PP_A_0.01_mgL', 'PP_A_0.1_mgL', 'PP_A_1_mgL', 'PP_A_10_mgL', 'PP_A_100_mgL'))

photobot_pp
```

```{r}
alpha_pp_aov <- aov(alpha ~ treatment,
  data = photobot_pp
)

summary(alpha_pp_aov)
```

```{r}
ggplot(photobot_pp) +
  aes(x = treatment, y = alpha, fill= treatment) +
  geom_boxplot() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
   labs(
    title = "Concentration vs. Alpha",
    x = "Leachate concentration (mg/L)",
    y = "Alpha"
  ) +
  scale_x_discrete(labels = c("Control ", "PPa 0.01", "PPa 0.1","PPa 1", "PPa 10", "PPa 100")) +  
  scale_fill_manual(values = c("seashell", "coral", "coral1","coral2","coral3", "coral4"))
```

# ETRm

```{r}
etrm_aov <- aov(ETRm ~ treatment,
  data = photobot_results
)

summary(etrm_aov)
```

```{r}
ggplot(photobot_results) +
  aes(x = treatment, y = ETRm, fill= treatment) +
  geom_boxplot() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
   labs(
    title = "Concentration vs. rETRm",
    x = "Leachate concentration (mg/L)",
    y = "rETRm"
  ) +
  scale_x_discrete(labels = c("Control ", "PPa 0.01", "PPa 0.1","PPa 1", "PPa 10", "PPa 100","PVC 0.01", "PVC 0.1", "PVC 1", "PVC 10", "PVC 100")) +  
  scale_fill_manual(values = c("seashell", "coral", "coral1","coral2","coral3", "coral4","dodgerblue","dodgerblue1","dodgerblue2", "dodgerblue3","dodgerblue4"))
```

# Ik

```{r}
ik_aov <- aov(Ik ~ treatment,
  data = photobot_results
)

summary(ik_aov)
```

```{r}
Ik <- ggplot(photobot_results) +
  aes(x = treatment, y = Ik, fill= treatment) +
  geom_boxplot() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
   labs(
    title = "Leachate concentration vs. Saturating Irradiance (Ik)",
    x = "Leachate concentration (mg/L)",
    y = "Saturating Irradiance (PAR umol/m2/s)"
  ) +
  scale_x_discrete(labels = c("Control ", "PPa 0.01", "PPa 0.1","PPa 1", "PPa 10", "PPa 100","PVC 0.01", "PVC 0.1", "PVC 1", "PVC 10", "PVC 100")) +  
  scale_fill_manual(values = c("seashell", "coral", "coral1","coral2","coral3", "coral4","dodgerblue","dodgerblue1","dodgerblue2", "dodgerblue3","dodgerblue4"))

Ik

```

What the hell is wrong with the control!?
